---
title: "French Study Abroad Project Analysis"
format: pdf
editor: visual
---

```{r}
# Load libraries
library(MASS)
library(tidyverse)
```

# Load Data

```{r}
# Load the preprocessed data
french_data <- read_csv("data/Processed_French_Data.csv",
                        show_col_types = FALSE)

french_data
```

# EDA

## Summary Statistics

The data has 58 observations and 14 variables that we will be analyzing. Twelve of these variables are predictors and two, Competence_Score and Intention_Score, are response variables.

```{r}
# Check for missing data
colSums(is.na(french_data))
```

The Screen_Time variable has 31 missing values. If we were to create a model using Screen_Time we would have to remove all of those observations, despite the fact that they are not missing in the other variables because a linear regression model cannot handle missing values.

Basic Summary Statistics for continuous variables

```{r}
continuous_variables <- c("Class_Level", "Screen_Time", "Interaction_Score",
                          "Speaking_Score", "Speak_Hours_Score", 
                          "Use_French_Score", "French_Hours_Score",
                          "Competence_Score", "Intention_Score")
summary(french_data[,continuous_variables])
```

It's important to note that Interaction_Score ranged from 0 to 40, Speaking_Score from 0 to 35, Speak_Hours_Score from 0 to 25, Use_French_Score from 0 to 28, French_Hours_Score from 0 to 20, Competence_Score from 0 to 80, and Intention_Score from 0 to 20. So all of these summary statistics look to be within their theoretical ranges. Screen_Time has a very large range from the minimum to the maximum, with the lowest being 2.4 hours per week on average, and the highest being about 47.8 hours per week on average. So that may require more looking into.

Counts for binary variables

```{r}
binary_variables <- c("Year_2024", "Year_2025", "Host_Family", "Phone_Language",
                      "Phone_Keyboard")

french_data |>
    group_by(Year_2024) |>
    summarize(count = n())
  
french_data |>
    group_by(Year_2025) |>
    summarize(count = n())

french_data |>
    group_by(Host_Family) |>
    summarize(count = n())

french_data |>
    group_by(Phone_Language) |>
    summarize(count = n())

french_data |>
    group_by(Phone_Keyboard) |>
    summarize(count = n())
```

Year_2024, Year_2025, Host_Family, and Phone_Language all have more 0 (no) entries than 1, whereas Phone_Keyboard has more 1s. Phone_Language is the only one that appears to have a roughly equal split, whereas the others are skewed.

Convert binary variables to factor.

```{r}
french_data <- french_data |>
  mutate(across(all_of(binary_variables), as.factor))
```

## Tables?

## Plots 

### Year

```{r}
french_data <- french_data |>
  mutate(Year = ifelse(Year_2024 == 1, "2024", 
                          ifelse(Year_2025 == 1, "2025", "2023"))) 

french_data |>
  group_by(Year)|>
  summarize(count = n())
```

```{r}
ggplot(french_data, aes(Year, fill = Year)) +
  geom_bar()
```

There's a difference in the number of students per year, but no clear pattern.

```{r}
ggplot(french_data, aes(Year, Competence_Score, color = Year)) +
  geom_boxplot()
```

For Year compared to Competence_Score, there doesn't seem to be any major difference in the overall distributions of the three years. The year 2025 has a slightly higher average Competence Score than 2023 or 2024.

```{r}
ggplot(french_data, aes(Year, Intention_Score, color = Year)) +
  geom_boxplot()
```

For Year vs Intention_Score, 2024 has a smaller IQR and range than 2023 and 2025, with a slightly higher average Intention_Score. Both 2023 and 2025 have very similar distributions.

### Class Level 

```{r}
ggplot(french_data, aes(Class_Level, fill = factor(Class_Level))) +
  geom_bar()
```

The Class_Level is somewhat right skewed with most students in the lowest levels and very few at the highest level.

```{r}
ggplot(french_data, aes(Class_Level, Competence_Score, 
                        color = factor(Class_Level))) +
  geom_boxplot()
```

There is a very clear correlation between increasing Class_Level and increasing perceived Competence_Score.

```{r}
ggplot(french_data, aes(Class_Level, Intention_Score, 
                        color = factor(Class_Level))) +
  geom_boxplot()
```

Similarly there is a fairly clear correlation between increasing Class_Level and increasing Intention_Score.

### Host Family

```{r}
ggplot(french_data, aes(Host_Family, fill = Host_Family)) +
  geom_bar()
```

```{r}
ggplot(french_data, aes(Host_Family, Competence_Score, color = Host_Family)) +
  geom_boxplot() +
    geom_jitter(width=0.2)
```

```{r}
ggplot(french_data, aes(Host_Family, Intention_Score, color = Host_Family)) +
  geom_boxplot()
```

### Phone Language

```{r}
ggplot(french_data, aes(Phone_Language, fill = Phone_Language)) +
  geom_bar()
```

```{r}
ggplot(french_data, aes(Phone_Language, Competence_Score, 
                        color = Phone_Language)) +
  geom_boxplot() +
  geom_jitter(width = 0.2)
```

```{r}
ggplot(french_data, aes(Phone_Language, Intention_Score, 
                        color = Phone_Language)) +
  geom_boxplot() +
  geom_jitter(width = 0.2)
```

### Phone Keyboard

```{r}
ggplot(french_data, aes(Phone_Keyboard, fill = Phone_Keyboard)) +
  geom_bar()
```

```{r}
ggplot(french_data, aes(Phone_Keyboard, Competence_Score, 
                        color = Phone_Keyboard)) +
  geom_boxplot() +
  geom_jitter(width = 0.2)
```

```{r}
ggplot(french_data, aes(Phone_Keyboard, Intention_Score, 
                        color = Phone_Keyboard)) +
  geom_boxplot() +
  geom_jitter(width = 0.2)
```

### Screen Time

```{r}
ggplot(french_data, aes(Screen_Time)) +
  geom_density(na.rm = TRUE)
```

```{r}
ggplot(french_data, aes(Screen_Time, Competence_Score)) +
  geom_point(na.rm=TRUE) +
  geom_smooth(method = 'lm', formula = 'y ~ x', na.rm=TRUE)
```

```{r}
ggplot(french_data, aes(Screen_Time, Intention_Score)) +
  geom_point(na.rm=TRUE) + 
  geom_smooth(method = "lm", formula = 'y ~ x', na.rm=TRUE)
```

### Interaction Score

```{r}
ggplot(french_data, aes(Interaction_Score)) +
  geom_density()
```

```{r}
ggplot(french_data, aes(Interaction_Score, Competence_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

```{r}
ggplot(french_data, aes(Interaction_Score, Intention_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

### Speaking Score

```{r}
ggplot(french_data, aes(Speaking_Score)) +
  geom_density()
```

```{r}
ggplot(french_data, aes(Speaking_Score, Competence_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

```{r}
ggplot(french_data, aes(Speaking_Score, Intention_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

### Speak Hour Score

```{r}
ggplot(french_data, aes(Speak_Hours_Score)) +
  geom_density()
```

```{r}
ggplot(french_data, aes(Speak_Hours_Score, Competence_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

```{r}
ggplot(french_data, aes(Speak_Hours_Score, Intention_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

### Use French Score

```{r}
ggplot(french_data, aes(Use_French_Score)) +
  geom_density()
```

```{r}
ggplot(french_data, aes(Use_French_Score, Competence_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

```{r}
ggplot(french_data, aes(Use_French_Score, Intention_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

### French Hours Score

```{r}
ggplot(french_data, aes(French_Hours_Score)) +
  geom_density()
```

```{r}
ggplot(french_data, aes(French_Hours_Score, Competence_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

```{r}
ggplot(french_data, aes(French_Hours_Score, Intention_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

### Competence and Intention Score

```{r}
ggplot(french_data, aes(Competence_Score)) +
  geom_density()
```

```{r}
ggplot(french_data, aes(Intention_Score)) +
  geom_density()
```

```{r}
ggplot(french_data, aes(Competence_Score, Intention_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

## Tables

### Table of Living Arrangements and what proportion of the total n each are, as well as what proportion of the total Competence Score they are

```{r}
french_data |> group_by(Host_Family) |> summarize(n = n(), comp = sum(Competence_Score)) |> mutate(prop_count = n/sum(n), prop_comp = comp/sum(comp))
```

Top row is Student dorms, and this row seems to be the majority of student living situations.

### Spread of data for Full dataset

```{r}
french_data |> lapply(summary)
```

### Average scores grouped by our Class level covariate

```{r}
french_data |> group_by(Class_Level) |> summarise(count=n(), avgcompscore = mean(Competence_Score), avgintscore = mean(Intention_Score), .groups = "drop")
```

Seems as though by class level, with 4 being the most difficult, competence scores do increase as one would expect.

## Graphs of variables

### Boxplot of screen time

```{r}
ggplot(french_data, aes(x= Screen_Time)) + geom_boxplot(fill = "orchid")
```

Looks to be somewhat skewed, also a couple outliers to the far right

### Histogram of Competence Score, normal curve overlayed

```{r}
ggplot(french_data, aes(x= Competence_Score)) + geom_histogram(aes(y = after_stat(density)), bins = 35) + stat_function(fun = dnorm, args = list(mean = mean(french_data$Competence_Score), sd = sd(french_data$Competence_Score)),color = "red")
```

Not awful, bi-modal skewed to the left a bit

### Histogram of Intention Score, normal curve overlayed

```{r}
ggplot(french_data, aes(x= Intention_Score)) + geom_histogram(aes(y = after_stat(density)), bins= 35) + stat_function(fun = dnorm, args = list(mean = mean(french_data$Intention_Score), sd = sd(french_data$Intention_Score)),color = "red")
```

Multi-modal histogram, not very symmetric

### Histogram of Interaction Score, normal curve overlayed

```{r}
ggplot(french_data, aes(x= Interaction_Score)) + geom_histogram(aes(y = after_stat(density)), bins= 35) + stat_function(fun = dnorm, args = list(mean = mean(french_data$Interaction_Score), sd = sd(french_data$Interaction_Score)),color = "red")
```

### Visualizing Sum of total competence score by Living Arrangement -\> French Speaking Home

```{r}
ggplot(french_data, aes(x = factor(Host_Family), y = Competence_Score, fill = factor(Class_Level))) +
  geom_bar(stat = "summary", fun = "mean") + scale_fill_brewer(palette = "Greens") +
  labs(x = "Lived with a Host Family (0 = No, 1 = Yes)", y = "Total Competence Score") + theme_classic()
```

```{r}
ggplot(french_data, aes(factor(Host_Family), Intention_Score, 
                        color=factor(Host_Family))) +
  geom_boxplot()
```

This bar plot shows that those that lived with a Host Family had a higher overall Competence score on average

### Plot of Interaction Score by Competence, colors are Class Level covariate

```{r}
ggplot(french_data, aes(x = Speaking_Score, y = Competence_Score, color = factor(Class_Level), shape = factor(Class_Level))) + geom_point() + labs(x = "Speaking Score", y = "Competence Score", color = "Class Level", shape = "Class Level")
```

This graph does not show a clear linear trend here

### Plot of Interaction Score by Intention, colors are Class Level covariate

```{r}
ggplot(french_data, aes(x = (Interaction_Score), y = Intention_Score, color = factor(Class_Level), shape = factor(Class_Level))) + geom_point()
```

Similar here, no obvious linear trend

### Screen Time Plotted by Competence Score

```{r}
ggplot(french_data, aes(x = (Screen_Time), y = Competence_Score, color = factor(Class_Level), shape = factor(Class_Level) )) + geom_point()
```

Similar results as those above

# Linear Models

The Screen Time variable had several issues with data quality and was not significant in either model, therefore we decided to remove it.

-   Screen Time does not have statistical significance in this model. However once again Immersion Score does. We also see that the global F test does not reject either, and with an adjusted R\^2 of .27, we cannot conclude much based on this model. Especially comparing this R\^2 to the R\^2 of the model not including screen time, this model seems to perform worse.

-   Also our client agreed that the variable didn't make sense as is and that we should remove it.

```{r}
competence_model <- lm(Competence_Score ~ Year_2024 + Year_2025 + Class_Level +
                         Host_Family + Phone_Language +
                         Phone_Keyboard + Interaction_Score + 
                         Speaking_Score + Speak_Hours_Score +
                         Use_French_Score + French_Hours_Score,
                       data = french_data)

summary(competence_model)
```

```{r}
intention_model <- lm(Intention_Score ~ Year_2024 + Year_2025 + 
                         Class_Level + Host_Family + 
                         Phone_Language + Phone_Keyboard +
                         Interaction_Score + Speaking_Score + Speak_Hours_Score +
                         Use_French_Score + French_Hours_Score,
                       data = french_data)

summary(intention_model)
```

# Model Checking using AIC

```{r}
stepAIC(competence_model, method = "backward")
```

```{r}
# After model checking
competence_model_reduced <- lm(Competence_Score ~ 
                         Class_Level +
                         Speaking_Score,
                       data = french_data)

summary(competence_model_reduced)
```

```{r}
# Competence Might Be slightly left skewed
plot(density(french_data$Competence_Score))
```

```{r}
# Residuals vs Fitted
plot(competence_model_reduced, which = 1)
```

Looks good, does not seem to have patterns or fan out or cluster too much. Some outliers.

Checking the outliers revealed that - for entry 45, the student had a higher perceived Competence score (82) than usual based on their fairly low Speaking Score (10) and Class Level (1) - For entry 16, the student had a lower perceived Competence score (29) than usual based on their fairly low Speaking Score (22) and Class Level (1) - For entry 19, the student had a lower perceived Competence score (37) than usual based on their fairly low Speaking Score (18) and Class Level (2)

# testing Competence Score for normality

```{r}
# QQ Plot
plot(competence_model_reduced, which = 2)
```

```{r}
shapiro.test(resid(competence_model_reduced))
```

```{r}
stepAIC(intention_model, method = "backward")
```

```{r}
# After model checking
intention_model_reduced <- lm(Intention_Score ~
                            Class_Level + 
                            Host_Family + 
                            Phone_Language + 
                            Interaction_Score +   
                            Speak_Hours_Score + 
                            Use_French_Score,
                       data = french_data)

summary(intention_model_reduced)
```

```{r}
# Intention Score is closer to a normal or at least symmetric distribution
plot(density(french_data$Intention_Score))
```

```{r}
# Residuals vs Fitted 
plot(intention_model_reduced, which = 1)
```

once again, we should check the noted outliers

-   for entry 24, the student had a higher Intention score (18) than usual based on their other entries in the model (class level 1, did not live with a french speaking family, did not have french as a phone language, above average interaction score etc)
-   For entry 3, the student had a lower Intention score (6) than usual based on their other entries in the model (class level 0, did not live with a french speaking family, did not french as a phone language, above average interaction score etc )
-   For entry 47, the student had a lower Intention score (14) than usual based on their other entries in the model (class level 4, lived with a french speaking family, had french as a phone language, above average interaction score etc)

# checking Intention Score for normality

```{r}
# QQ Plot
plot(intention_model_reduced, which = 2)
```

Looks fairly normally distributed

```{r}
shapiro.test(resid(intention_model_reduced))
```

pval \> 0.05, we can assume normality

Let's look at plots of Interaction_Score, Speaking_Score, Speak_Hours_Score, Use_French_Score, and French_Hours_Score vs Competence_Score and Intention_Score

```{r}
# Something interesting I found is that Competence and Intention have 
# a moderately positive correlation 
ggplot(french_data, aes(Competence_Score, Intention_Score)) +
  geom_point() + 
  geom_smooth(method="lm", formula = y ~ x)
```

```{r}
french_data |>
  dplyr::select(everything(), -Screen_Time) |>
  cor(use = "complete.obs")
```

Ignoring the variables that are a composite of other variables, none of the variables are too highly correlated to the point that multicollinearity would be an issue. We can see that Read_French_Score, Write_French_Score, Listen_French_Score, and Speak_French_Score all have moderate to strong positive correlations with one another. Interestingly enough Competence_Score and Intention_Score almost had only positive correlations with all of the other variables, albeit many of those correlations are weak.

Might need to standardize the variables first, but for now it's probably fine
