---
title: "French Study Abroad Project Analysis"
format: pdf
editor: visual
---

```{r}
# Load libraries
library(MASS)
library(tidyverse)
library(car)
```

# Load Data

```{r}
# Load the preprocessed data
french_data <- read_csv("data/Processed_French_Data.csv",
                        show_col_types = FALSE)

french_data
```

# EDA

## Summary Statistics

The data has 58 observations and 14 variables that we will be analyzing. Twelve of these variables are predictors and two, Competence_Score and Intention_Score, are response variables.

```{r}
# Check for missing data
colSums(is.na(french_data))
```

The Screen_Time variable has 31 missing values. If we were to create a model using Screen_Time we would have to remove all of these observations, despite the fact that they are not missing in the other variables because a linear regression model cannot handle missing values.

Basic Summary Statistics for continuous variables

```{r}
continuous_variables <- c("Class_Level", "Screen_Time", "Interaction_Score",
                          "Speaking_Score", "Speak_Hours_Score", 
                          "Use_French_Score", "French_Hours_Score",
                          "Competence_Score", "Intention_Score")
summary(french_data[,continuous_variables])
```

It's important to note that Interaction_Score ranged from 0 to 40, Speaking_Score from 0 to 35, Speak_Hours_Score from 0 to 25, Use_French_Score from 0 to 28, French_Hours_Score from 0 to 20, Competence_Score from 0 to 80, and Intention_Score from 0 to 20. So all of these summary statistics look to be within their theoretical ranges. Screen_Time has a very large range from the minimum to the maximum, with the lowest being 2.4 hours per week on average, and the highest being about 47.8 hours per week on average. So that may require more looking into.

Counts for binary variables

```{r}
binary_variables <- c("Year_2024", "Year_2025", "Host_Family", "Phone_Language",
                      "Phone_Keyboard")

french_data |>
    group_by(Year_2024) |>
    summarize(count = n())
  
french_data |>
    group_by(Year_2025) |>
    summarize(count = n())

french_data |>
    group_by(Host_Family) |>
    summarize(count = n())

french_data |>
    group_by(Phone_Language) |>
    summarize(count = n())

french_data |>
    group_by(Phone_Keyboard) |>
    summarize(count = n())
```

Year_2024, Year_2025, Host_Family, and Phone_Language all have more 0 (no) entries than 1, whereas Phone_Keyboard has more 1s. Phone_Language is the only one that appears to have a roughly equal split, whereas the others are skewed.

Convert binary variables to factor.

```{r}
# Save original numeric data for calculating correlations
french_data_corr <- french_data

# Convert binary variables to factor
french_data <- french_data |>
  mutate(across(all_of(binary_variables), as.factor))
```

## Plots

### Year

```{r}
french_data <- french_data |>
  mutate(Year = ifelse(Year_2024 == 1, "2024", 
                          ifelse(Year_2025 == 1, "2025", "2023"))) 

french_data |>
  group_by(Year)|>
  summarize(count = n())
```

```{r}
ggplot(french_data, aes(Year, fill = Year)) +
  geom_bar()
```

There's a difference in the number of students per year, but no clear pattern. 2023 has the most, then 2025, and finally 2024.

```{r}
ggplot(french_data, aes(Year, Competence_Score, color = Year)) +
  geom_boxplot()
```

For Year compared to Competence_Score, there doesn't seem to be any major difference in the overall distributions of the three years. The year 2025 has a higher average Competence Score than 2023 or 2024.

```{r}
ggplot(french_data, aes(Year, Intention_Score, color = Year)) +
  geom_boxplot()
```

For Year vs Intention_Score, 2024 has a smaller IQR and range than 2023 and 2025, as well as a slightly higher average Intention_Score. Both 2023 and 2025 have very similar distributions.

### Class Level

```{r}
french_data |> 
  group_by(Class_Level) |> 
  summarise(count=n(), avgcompscore = mean(Competence_Score), 
            avgintscore = mean(Intention_Score), .groups = "drop")
```

Seems as though by class level, with 4 being the most difficult, competence scores do increase as one would expect.

```{r}
ggplot(french_data, aes(Class_Level, fill = factor(Class_Level))) +
  geom_bar()
```

The Class_Level is somewhat right skewed with most students in the lowest levels and very few at the highest level.

```{r}
ggplot(french_data, aes(Class_Level, Competence_Score, 
                        color = factor(Class_Level))) +
  geom_boxplot()
```

There is a very clear correlation between increasing Class_Level and increasing perceived Competence_Score.

```{r}
ggplot(french_data, aes(Class_Level, Intention_Score, 
                        color = factor(Class_Level))) +
  geom_boxplot()
```

Similarly there is a fairly clear correlation between increasing Class_Level and increasing Intention_Score on average.

### Host Family

```{r}
french_data |> 
  group_by(Host_Family) |> 
  summarize(n = n(), comp = sum(Competence_Score)) |> 
  mutate(prop_count = n/sum(n), prop_comp = comp/sum(comp))
```

Top row is Student housing, and this row seems to be the majority of student living situations.

```{r}
ggplot(french_data, aes(Host_Family, fill = Host_Family)) +
  geom_bar()
```

```{r}
ggplot(french_data, aes(x = Host_Family, y = Competence_Score, 
                        fill = factor(Class_Level))) +
  geom_bar(stat = "summary", fun = "mean") + 
  scale_fill_brewer(palette = "Greens") +
  labs(x = "Lived with a Host Family (0 = No, 1 = Yes)", 
       y = "Total Competence Score") + 
  theme_classic()
```

This bar plot shows that those that lived with a Host Family had a higher overall Competence score on average

(This is a weird plot to me because you're looking at the total competence across two groups, but the groups have different numbers of students in them. I guess it's interesting that students in class level 4 are only living with a host family.)

```{r}
ggplot(french_data, aes(Host_Family, Competence_Score, color = Host_Family)) +
  geom_boxplot() +
    geom_jitter(width=0.2)
```

On average the students who lived with a host family had a slightly higher competence score.

```{r}
ggplot(french_data, aes(Host_Family, Intention_Score, color = Host_Family)) +
  geom_boxplot()
```

Similarly, on average the students who lived with a host family had a slightly higher intention score.

### Phone Language

```{r}
ggplot(french_data, aes(Phone_Language, fill = Phone_Language)) +
  geom_bar()
```

Slightly more students did not have French as a language on their phone.

```{r}
ggplot(french_data, aes(Phone_Language, Competence_Score, 
                        color = Phone_Language)) +
  geom_boxplot() +
  geom_jitter(width = 0.2)
```

The students who did have French as a language on their phone had a higher competence score on average.

```{r}
ggplot(french_data, aes(Phone_Language, Intention_Score, 
                        color = Phone_Language)) +
  geom_boxplot() +
  geom_jitter(width = 0.2)
```

Similarly, the students who did have French as a language on their phone had a higher intention score on average.

### Phone Keyboard

```{r}
ggplot(french_data, aes(Phone_Keyboard, fill = Phone_Keyboard)) +
  geom_bar()
```

There are far more students who have a French keyboard on their phone.

```{r}
ggplot(french_data, aes(Phone_Keyboard, Competence_Score, 
                        color = Phone_Keyboard)) +
  geom_boxplot() +
  geom_jitter(width = 0.2)
```

While the IQR and range are wider for the French keyboard group as you might expect from having more data, the mean competence score for both groups are about the same.

```{r}
ggplot(french_data, aes(Phone_Keyboard, Intention_Score, 
                        color = Phone_Keyboard)) +
  geom_boxplot() +
  geom_jitter(width = 0.2)
```

The mean intention score for the French keyboard group is much higher than the group that didn't use a French keyboard.

### Screen Time

```{r}
ggplot(french_data, aes(Screen_Time)) +
  geom_density(na.rm = TRUE) + 
  stat_function(fun = dnorm, 
                args = list(mean = mean(french_data$Screen_Time, na.rm=TRUE), 
                            sd = sd(french_data$Screen_Time, na.rm=TRUE)),
                color = "red")
```

It looks like Screen_Time may have a left skewed distribution with a higher mean than median. It's also possible that the distribution is bimodal which may indicate that students are answering the question in different ways.

```{r}
ggplot(french_data, aes(x= Screen_Time)) + 
  geom_boxplot(fill = "orchid", na.rm = TRUE)
```

The boxplot makes it more clear that the reason for the higher mean compared to the median may be due to a few outliers on the far right.

```{r}
ggplot(french_data, aes(Screen_Time, Competence_Score)) +
  geom_point(na.rm=TRUE) +
  geom_smooth(method = 'lm', formula = 'y ~ x', na.rm=TRUE)
```

A perfectly flat model is possible in the 95% confidence bands, which indicates there may be no linear relationship between Screen_Time and Competence_Score.

```{r}
ggplot(french_data, aes(x = Screen_Time, y = Competence_Score, 
                        color = factor(Class_Level), 
                        shape = factor(Class_Level) )) + 
  geom_point(na.rm = TRUE)
```

Similar results as those above - no clear linear trend.

```{r}
ggplot(french_data, aes(Screen_Time, Intention_Score)) +
  geom_point(na.rm=TRUE) + 
  geom_smooth(method = "lm", formula = 'y ~ x', na.rm=TRUE)
```

Once again, a perfectly flat line is possible in the 95% confidence bands, which indicates that there may not be a linear relationship between Screen_Time and Intention_Score.

Due to the missing values, bimodal nature of the distribution, along with the outliers and the potential for there being no linear relationship with Competence_Score and Intention_Score, it seems likely that Screen_Time will be removed from the model.

### Interaction Score

```{r}
ggplot(french_data, aes(Interaction_Score)) +
  geom_density() + 
  stat_function(fun = dnorm, 
                args = list(mean = mean(french_data$Interaction_Score), 
                            sd = sd(french_data$Interaction_Score)),color = "red")
```

The Interaction_Score appears to follow the normal distribution quite well given the amount of data in the sample.

```{r}
ggplot(french_data, aes(Interaction_Score, Competence_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

There appears to be a slight positive linear relationship between Interaction_Score and Competence_Score, however a perfectly flat line may be possible in the 95% confidence bands, which would indicate no relationship.

```{r}
ggplot(french_data, aes(x = Interaction_Score, y = Competence_Score, 
                        color = factor(Class_Level), 
                        shape = factor(Class_Level))) + 
  geom_point() + 
  labs(x = "Speaking Score", y = "Competence Score", 
       color = "Class Level", shape = "Class Level")
```

This graph does not show a clear linear trend here

```{r}
ggplot(french_data, aes(Interaction_Score, Intention_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

This plot indicates that there is a positive linear relationship between Interaction_Score and Intention_Score.

```{r}
ggplot(french_data, aes(x = Interaction_Score, y = Intention_Score, 
                        color = factor(Class_Level), 
                        shape = factor(Class_Level))) + 
  geom_point()
```

Similar here, no obvious linear trend

### Speaking Score

```{r}
ggplot(french_data, aes(Speaking_Score)) +
  geom_density() + 
  stat_function(fun = dnorm, 
                args = list(mean = mean(french_data$Speaking_Score), 
                            sd = sd(french_data$Speaking_Score)),color = "red")
```

Given the amount of data in the sample, Speaking_Score is roughly following a normal distribution, however it's possible that it deviates slightly. From the summary statistics above, the mean and median are very close to one another.

```{r}
ggplot(french_data, aes(Speaking_Score, Competence_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

There appears to be a positive linear relationship between Speaking_Score and Competence_Score.

```{r}
ggplot(french_data, aes(Speaking_Score, Intention_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

There also appears to be a positive linear relationship between Speaking_Score and Intention_Score, but it's possible that a flat line is within the 95% confidence bands which would indicate no relationship.

### Speak Hour Score

```{r}
ggplot(french_data, aes(Speak_Hours_Score)) +
  geom_density() + 
  stat_function(fun = dnorm, 
                args = list(mean = mean(french_data$Speak_Hours_Score), 
                            sd = sd(french_data$Speak_Hours_Score)),color = "red")
```

The Speak_Hours_Score appears to be left skewed and from the summary statistics it has a much higher mean than median.

```{r}
ggplot(french_data, aes(Speak_Hours_Score, Competence_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

There is a positive linear relationship between Speak_Hours_Score and Competence_Score.

```{r}
ggplot(french_data, aes(Speak_Hours_Score, Intention_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

Similarly, there is a positive linear relationship between Speak_Hours_Score and Intention_Score.

### Use French Score

```{r}
ggplot(french_data, aes(Use_French_Score)) +
  geom_density() + 
  stat_function(fun = dnorm, 
                args = list(mean = mean(french_data$Use_French_Score), 
                            sd = sd(french_data$Use_French_Score)),color = "red")
```

The density shows that Use_French_Score may be slightly left skewed, which is backed up by the summary statistics which show that the mean is higher than the median. However, given the relatively small sample size, it's possible this is just due to variance.

```{r}
ggplot(french_data, aes(Use_French_Score, Competence_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

There is a positive linear relationship between Use_French_Score and Competence_Score, however a flat line may be possible in the 95% confidence bands, which would indicate there may not be a linear relationship.

```{r}
ggplot(french_data, aes(Use_French_Score, Intention_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

A flat line between Use_French_Score and Intention_Score is possible in the 95% confidence bands, which indicates there may not be a linear relationship.

### French Hours Score

```{r}
ggplot(french_data, aes(French_Hours_Score)) +
  geom_density() + 
  stat_function(fun = dnorm, 
                args = list(mean = mean(french_data$French_Hours_Score), 
                            sd = sd(french_data$French_Hours_Score)),color = "red")
```

The French_Hours_Score is left skewed which is also backed up by the summary statistics showing that it has a much higher mean than median.

```{r}
ggplot(french_data, aes(French_Hours_Score, Competence_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

A flat line is possible in the 95% confidence bands, therefore there may not be a linear relationship between French_Hours_Score and Competence_Score.

```{r}
ggplot(french_data, aes(French_Hours_Score, Intention_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

A flat line is also possible in the 95% confidence bands here, therefore there may not be a linear relationship between French_Hours_Score and Intention_Score.

### Competence and Intention Score

```{r}
ggplot(french_data, aes(Competence_Score)) +
  geom_density() +
  stat_function(fun = dnorm, 
                args = list(mean = mean(french_data$Competence_Score), 
                            sd = sd(french_data$Competence_Score)),color = "red")
```

Given the sample size, the Competence_Score roughly follows the normal distribution. Additionally, looking at the summary statistics, the mean and median were very close.

```{r}
ggplot(french_data, aes(Intention_Score)) +
  geom_density() +
  stat_function(fun = dnorm, 
                args = list(mean = mean(french_data$Intention_Score), 
                            sd = sd(french_data$Intention_Score)),color = "red")
```

Similarly, given the sample size, the Intention_Score also roughly follows a normal distribution. It's mean and median were also very close to one another.

```{r}
ggplot(french_data, aes(Competence_Score, Intention_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

There is a very clear positive linear relationship between Competence_Score and Intention_Score.

### Correlations

```{r}
# Look at the correlations with Screen_Time separately since it removes NA values
ST_corr <- french_data_corr |>
  select(Year_2024, Year_2025, Class_Level, Host_Family, Phone_Language,
         Phone_Keyboard, Screen_Time, Interaction_Score, Speaking_Score,
         Speak_Hours_Score, Use_French_Score, French_Hours_Score,
         Competence_Score, Intention_Score) |>
  cor(use = "complete.obs") 

ST_corr[,"Screen_Time"]
```

As can be seen, the Screen_Time variable has low correlations with all of the other variables, including weak correlations with both Competence_Score and Intention_Score. This is more evidence that Screen_Time may need to be removed from the data set.

```{r}
# Correlations for the rest of the data
french_data_corr |>
  select(Year_2024, Year_2025, Class_Level, Host_Family, Phone_Language,
         Phone_Keyboard, Interaction_Score, Speaking_Score,
         Speak_Hours_Score, Use_French_Score, French_Hours_Score,
         Competence_Score, Intention_Score) |>
  cor(use = "complete.obs")
```

Correlation strength will be defined as:

-   0 to abs(0.2) - Very weak or no association

-   abs(0.2) to abs(0.4) - Weak association

-   abs(0.4) to abs(0.6) - Moderation association

-   abs(0.6) to abs(0.8) - Strong association

-   abs(0.8) to abs(1.0) - Perfect or very strong association

The variables with nearly zero or very weak linear correlation with Competence_Score were: Year_2024, Year_2025, Phone_Keyboard, and French_Hours_Score. The ones with weak linear correlation were: Host_Family, Interaction_Score, Speaking_Score, Speak_Hours_Score, and Use_French_Score. The variables with moderate linear correlation were: Phone_Language and Intention_Score. Finally the only variable with a strong association was Class_Level.

The variables with nearly zero or very weak linear correlation with Intention_Score were: Year_2024, Year_2025, Phone_Keyboard, and Use_French_Score. The variables with weak linear correlation were: Host_Family, Interaction_Score, Speaking_Score, Speak_Hours Score, and French_Hours_Score. The variables with moderate linear correlation were: Class_Level, Phone_Language, and Competence_Score.

Speaking_Score has a strong association with Use_French_Score. Speak_Hours_Score has a strong association with Use_French_Score and a very strong association with French_Hours_Score. Finally, Use_French_Score also has a strong association with French_Hours Score. It's possible that this may cause some of these variables to have too high of multicollinearity that will affect the model fit and whether they are statistically significant.

```{r}
ggplot(french_data, aes(Speaking_Score, Use_French_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

```{r}
ggplot(french_data, aes(Speak_Hours_Score, Use_French_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

```{r}
ggplot(french_data, aes(Speak_Hours_Score, French_Hours_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

```{r}
ggplot(french_data, aes(Use_French_Score, French_Hours_Score)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = 'y ~ x')
```

# Linear Models

The Screen_Time variable had several issues with data quality, and along with the various details found above, our client determined that it had too many issues. Therefore, the variable was not included in the model.

Might need to standardize the variables first, but for now it's probably fine

```{r}
competence_model <- lm(Competence_Score ~ Year_2024 + Year_2025 + Class_Level +
                         Host_Family + Phone_Language +
                         Phone_Keyboard + Interaction_Score + 
                         Speaking_Score + Speak_Hours_Score +
                         Use_French_Score + French_Hours_Score,
                       data = french_data)

summary(competence_model)
```

```{r}
# Check the variance inflation factor 
vif(competence_model)
```

A VIF of 1 - 5 indicates low to moderate multicollinearity, and since none of the VIFs were above 5, the model fit is likely not being affected by it.

```{r}
intention_model <- lm(Intention_Score ~ Year_2024 + Year_2025 + 
                         Class_Level + Host_Family + 
                         Phone_Language + Phone_Keyboard +
                         Interaction_Score + Speaking_Score + Speak_Hours_Score +
                         Use_French_Score + French_Hours_Score,
                       data = french_data)

summary(intention_model)
```

```{r}
# Check the variance inflation factor 
vif(intention_model)
```

Similar to above, none of the VIFs were above 5, therefore the model fit is likely not being affected by multicollinearity.

## Feature Selection using BIC

BIC is the Bayesian Information Criterion which can be used to select which model is fitting the data better based on which has the lowest BIC. It has a stronger penalty for models with more parameters. The equation is BIC = -2\*ln(Likelihood) + k\*ln(n), where k is the number of parameters in the model and n is the number of observations in the dataset. We decided to use BIC because we wanted to obtain a more parsimonious model. The feature selection method uses a backward elimination approach which removes one variable at a time and chooses the model that reduced the BIC the most. Backward elimination is the default method for performing feature selection in this way. Using forward or both as the method for selection provided the same results for both models.

## Competence Model

```{r}
# Feature selection using backwards elimation and BIC
# k is what defines the model the be BIC even though the output says AIC
competence_model_reduced <- step(competence_model, method = "backward",
                                 k = log(nrow(competence_model$model)))
summary(competence_model_reduced)
```

Using backward elimination based on BIC, we get a final reduced model with Class_Level and Speaking_Score. However, this model rejected normality of the residuals in the Shapiro-Wilks test, which means the model assumptions were not met and the model was not valid. The BIC difference between the final model and the one above it was only 1.12, which is very small. Therefore we chose to use the model with Interaction_Score.

```{r}
# Reverting to the second best BIC model to meet model assumptions
competence_model_reduced <- lm(Competence_Score ~
                                Class_Level + 
                                Interaction_Score +
                                Speaking_Score,
                              data = french_data)

summary(competence_model_reduced)
```

In this model the Class_Level and Speaking_Score variables are statistically significant at a significance level of 0.05. Even though the Interaction_Score is not statistically significant at a significance level of 0.05, including it in the model allows the model to meet model assumptions.

```{r}
# Residuals vs Fitted
plot(competence_model_reduced, which = 1)
```

Generally the residuals are spread out evenly across the fitted values, showing that there is roughly equal variance of the residuals. However, the three outliers identified by the plot are observations 16, 19, and 49.

```{r}
# QQ Plot
plot(competence_model_reduced, which = 2)
```

The standardized residuals only slightly deviate from the theoretical normal quantiles at the ends, indicating that the residuals are likely normally distributed.

```{r}
shapiro.test(resid(competence_model_reduced))
```

The Shapiro-Wilk test tests the null hypothesis that a sample comes from a normally distributed population. With a p-value of 0.1359, this model fails to reject the null hypothesis at a significance level of 0.05. Therefore, there is evidence to support that the sample does come from a normally distributed population.

```{r}
# Once again we can check for multicollinearity
vif(competence_model_reduced)
```

All of the VIFs are very low which indicates that there is very little multicollinearity.

Looking at the outliers from the plots above, observations 16 and 19 have large negative residuals and 49 has a large positive one. A residual is the actual value - the predicted one. So a negative residual means the model overestimated, and a positive means it underestimated.

```{r}
french_data |>
  mutate(Observation = seq(1:nrow(french_data))) |>
  select(Class_Level, Interaction_Score, Speaking_Score, 
         Competence_Score, Observation) |>
  arrange(Competence_Score)
```

As can be seen from the data, observations 16 and 19 had relatively low Competence_Scores given their Class_Level, Interaction_Score, and Speaking_Score. On the other hand observation 49 has a surprisingly high Competence_Score given it's relatively low Class_Level, Interaction_Score, and Speaking_Score.

```{r}
cooksdc <- cooks.distance(competence_model_reduced)
plot(cooksdc, xlab = "Observation Number", ylab = "Cook's Distance")
abline(h = 4/nrow(french_data), col = "red", lty = 2)
inflabc <- which(cooksdc > 4/nrow(french_data))
text(inflabc, cooksdc[inflabc], labels = inflabc, pos=2)
```

Looking at Cook's Distance we see that 32 and 49 are both influential values. Observation 49 was discussed above, but as can be seen observation 32 is the value with the lowest Competence_Score as well as the Speaking_Score being 0.

## Intention Model

```{r}
# Feature selection using backwards elimation and BIC
# k is what defines the model the be BIC even though the output says AIC
intention_model_reduced <- step(intention_model, method = "backwards",
                                k = log(nrow(competence_model$model)))
summary(intention_model_reduced)
```

The final reduced model included Class_Level, Phone_Language, and Interaction_Score. All of these variables are statistically significant at a significance level of 0.05.

```{r}
# Residuals vs Fitted 
plot(intention_model_reduced, which = 1)
```

The residuals appear to be roughly evenly spaced across the fitted values, indicating there is equal variance of the residuals. The three outliers identified by the plot are observations 3, 22, and 35.

```{r}
# QQ Plot
plot(intention_model_reduced, which = 2)
```

The standardized residuals appear to align roughly with the theoretical normal quantiles, indicating that the residuals are normally distributed.

```{r}
shapiro.test(resid(intention_model_reduced))
```

With a p-value of 0.1943, this model fails to reject the null hypothesis at a significance level of 0.05. Therefore, there is evidence to support that the sample does come from a normally distributed population.

```{r}
# Once again we can check for multicollinearity
vif(intention_model_reduced)
```

All of the VIFs are very low which indicates that there is very little multicollinearity.

Looking at the outliers from the plots above, observations 3 and 35 have large negative residuals and 22 has a large positive one. A residual is the actual value - the predicted one. So a negative residual means the model overestimated, and a positive means it underestimated.

```{r}
french_data |>
  mutate(Observation = seq(1:nrow(french_data))) |>
  select(Class_Level, Phone_Language, Interaction_Score, 
         Intention_Score, Observation) |>
  arrange(Intention_Score)
```

While it's somewhat hard to see it appears that the Intention_Score for observation 3 and 35 were lower than the model expected given their Class_Level, Phone_Language, and Interaction_Score combination. With observation 22 it is easier to see that it has a higher Intention_Score than the model expected given a Class_Level of 1, as well as the Phone_Language and Interaction_Score variables.

```{r}
cooksdi <- cooks.distance(intention_model_reduced)
plot(cooksdi, xlab = "Observation Number", ylab = "Cook's Distance")
abline(h = 4/nrow(french_data), col = "red", lty = 2)
inflabi <- which(cooksdi > 4/nrow(french_data))
text(inflabi, cooksdi[inflabi], labels = inflabi, pos=2)
```

Observation 3 is an influential point and it was described above.
