---
title: "French Study Abroad Project Analysis"
format: pdf
editor: visual
---

# Load Data

```{r}
# Load the preprocessed data
french_data <- read_csv("data/processed_french_data.csv")

french_data
```

# EDA

## Tables

### Table of Living Arrangements and what proportion of the total n each are, as well as what proportion of the total Competence Score they are

```{r}
french_data |> group_by(Host_Family) |> summarize(n = n(), comp = sum(Competence_Score)) |> mutate(prop_count = n/sum(n), prop_comp = comp/sum(comp))
```

Top row is Student dorms, and this row seems to be the majority of student living situations.

### Spread of data for Full dataset that does not include Screen Time

```{r}
french_data |> lapply(summary)
```

### Average scores grouped by our Class level covariate

```{r}
french_data |> group_by(Class_Level) |> summarise(count=n(), avgcompscore = mean(Competence_Score), avgintscore = mean(Continue_French_Score), .groups = "drop")
```

Seems as though by class level, with 4 being the most difficult, competence scores do increase as one would expect.

## Graphs of variables

### Boxplot of screen time

```{r}
ggplot(french_data, aes(x= Screen_Time)) + geom_boxplot(fill = "orchid")
```

### Histogram of Competence Score, normal curve overlayed

```{r}
ggplot(french_data, aes(x= Competence_Score)) + geom_histogram(aes(y = after_stat(density)), bins = 35) + stat_function(fun = dnorm, args = list(mean = mean(french_data$Competence_Score), sd = sd(french_data$Competence_Score)),color = "red")
```

### Histogram of Intention Score, normal curve overlayed

```{r}
ggplot(french_data, aes(x= Continue_French_Score)) + geom_histogram(aes(y = after_stat(density)), bins= 35) + stat_function(fun = dnorm, args = list(mean = mean(french_data$Continue_French_Score), sd = sd(french_data$Continue_French_Score)),color = "red")
```

### Histogram of Interaction Score, normal curve overlayed

```{r}
ggplot(french_data, aes(x= Interaction_Score)) + geom_histogram(aes(y = after_stat(density)), bins= 35) + stat_function(fun = dnorm, args = list(mean = mean(french_data$Interaction_Score), sd = sd(french_data$Interaction_Score)),color = "red")
```

### Visualizing Sum of total competence score by Living Arrangement -\> French Speaking Home

```{r}
ggplot(french_data, aes(x = factor(Host_Family), y = Competence_Score, fill = factor(Class_Level))) +
  geom_bar(stat = "summary", fun = "mean") + scale_fill_brewer(palette = "Reds") +
  labs(x = "Lived with a Host Family (0 = No, 1 = Yes)", y = "Total Competence Score") + theme_classic()
```

This bar plot shows that those that lived with a Host Family had a higher overall Competence score on average

### Plot of Immersion Score by Competence, colors are Class Level covariate

```{r}
ggplot(french_data, aes(x = (Interaction_Score), y = Competence_Score, color = factor(Class_Level))) + geom_point() + geom_abline()
```

This graph does not show a clear linear trend here

### Plot of Immersion Score by Intention, colors are Class Level covariate

```{r}
ggplot(french_data, aes(x = (Interaction_Score), y = Continue_French_Score, color = factor(Class_Level))) + geom_point()
```

Similar here, no obvious linear trend

### Screen Time Plotted by Competence Score

```{r}
ggplot(french_data, aes(x = (Screen_Time), y = Competence_Score, color = factor(Class_Level))) + geom_point()
```

Similar results as those above

# Linear Models

The Screen Time variable had several issues with data quality and was not significant in either model, therefore we decided to remove it.

-   Screen Time does not have statistical significance in this model. However once again Immersion Score does. We also see that the global F test does not reject either, and with an adjusted R\^2 of .27, we cannot conclude much based on this model. Especially comparing this R\^2 to the R\^2 of the model not including screen time, this model seems to perform worse.

-   Also our client agreed that the variable didn't make sense as is and that we should remove it.

```{r}
competence_model2 <- lm(Competence_Score ~ Year_2024 + Year_2025 + Class_Level +
                         Host_Family + Phone_Language +
                         Phone_Keyboard + Interaction_Score + 
                         Speaking_Score + Speak_Hours_Score +
                         Use_French_Score + French_Hours_Score,
                       data = french_data)

summary(competence_model2)
```

```{r}
intention_model2 <- lm(Continue_French_Score ~ Year_2024 + Year_2025 + 
                         Class_Level + Host_Family + 
                         Phone_Language + Phone_Keyboard +
                         Interaction_Score + Speaking_Score + Speak_Hours_Score +
                         Use_French_Score + French_Hours_Score,
                       data = french_data)

summary(intention_model2)
```

```{r}
# Removing the least significant variable until the remaining are significant
competence_model4 <- lm(Competence_Score ~ 
                         Class_Level +
                         Interaction_Score + 
                         Speaking_Score,
                       data = french_data)

summary(competence_model4)
```

```{r}
# Competence Might Be slightly left skewed
plot(density(french_data$Competence_Score))
```

```{r}
# Residuals vs Fitted
plot(competence_model4, which = 1)
```

```{r}
# Holly's version of Fitted Values vs Residual Plot
ggplot(data.frame(fitted = fitted(Comp_model), resid = residC), aes(x = fitted, y = residC)) +
  geom_point()
```

Looks good, does not seem to have patterns or fan out or cluster too much

```{r}
# QQ Plot - This is still probably fine and the assumption is still met
# Though the far left two points may be outliers
plot(competence_model4, which = 2)
```

```{r}
# Holly's Version of QQ Plot
residI <- residuals(Int_model_IM)
ggplot(data.frame(residI), aes(sample = residI)) + stat_qq() + stat_qq_line()
```

```{r}
# Removing the least significant variable until the remaining are significant
# alpha = 0.5 and 0.1
intention_model4 <- lm(Continue_French_Score ~
                         Class_Level +
                         Phone_Language +
                         Interaction_Score,
                       data = french_data)

summary(intention_model4)
```

```{r}
# Intention Score is closer to a normal or at least symmetric distribution
plot(density(french_data$Continue_French_Score))
```

```{r}
# Residuals vs Fitted 
plot(intention_model4, which = 1)
```

```{r}
# QQ Plot
plot(intention_model4, which = 2)
```

Let's look at plots of Interaction_Score, Speaking_Score, Speak_Hours_Score, Use_French_Score, and French_Hours_Score vs Competence_Score and Continue_French_Score

```{r}
# Something interesting I found is that Competence and Intention have 
# a moderately positive correlation 
ggplot(french_data, aes(Competence_Score, Continue_French_Score)) +
  geom_point() + 
  geom_smooth(method="lm", formula = y ~ x)
```

```{r}
french_data |>
  select(everything(), -Screen_Time) |>
  cor(use = "complete.obs")
```

Ignoring the variables that are a composite of other variables, none of the variables are too highly correlated to the point that multicollinearity would be an issue. We can see that Read_French_Score, Write_French_Score, Listen_French_Score, and Speak_French_Score all have moderate to strong positive correlations with one another. Interestingly enough Competence_Score and Continue_French_Score almost had only positive correlations with all of the other variables, albeit many of those correlations are weak.

Might need to standardize the variables first, but for now it's probably fine
