---
title: "French Study Abroad Project Analysis"
format: pdf
editor: visual
---

# Linear Models
# Model that does not include Screen Time for y = Competence Score
```{r}
Comp_model <- lm(Competence_Score ~ Immersion_Score + Host_French + Host_English + Phone_Language + Phone_Keyboard + Class_Level + Year_2024 + Year_2025, data = french_data)
summary(Comp_model)
```
Seeing that Immersion Score has statistical significance, it is of interest to break down into its categories and test again

```{r}
Comp_model_IM <- lm(Competence_Score ~ Interaction_Score + Speaking_Score + Speak_Hours_Score + Use_French_Score +]] French_Hours_Score + Host_French + Host_English + Phone_Language + Phone_Keyboard + Class_Level + Year_2024 + Year_2025, data = french_data)
summary(Comp_model_IM)
```
None of them stand out individually 

# Model that does not include Screen Time for y = Continue French Score
```{r}
Int_model <- lm(Continue_French_Score ~ Immersion_Score + Host_French + Host_English + Phone_Language + Phone_Keyboard + Class_Level + Year_2024 + Year_2025, data = french_data)
summary(Int_model)
```
Again the overall Immersion score is statistically signficant, as well as what language they have their phone set to 

```{r}
Int_model_IM <- lm(Continue_French_Score ~ Interaction_Score + Speaking_Score + Speak_Hours_Score + Use_French_Score + French_Hours_Score + Host_French + Host_English + Phone_Language + Phone_Keyboard + Class_Level + Year_2024 + Year_2025, data = french_data)
summary(Int_model_IM)
```
Here, the Interaction Score is significant within Immersion Score. 

# Model that includes Screen Time for y = Competence Score
```{r}
STComp_model <- lm(Competence_Score ~ Immersion_Score + Host_French + Host_English + Phone_Language + Phone_Keyboard + Class_Level + Year_2024 + Year_2025 + Screen_Time, data = WorkingSTData)
summary(STComp_model)
```
Screen Time does not have statistical significance in this model. However once again Immersion Score does. We also see that the global F test does not reject either, and with an adjusted R^2 of .27, we cannot conclude much based on this model. Especially comparing this R^2 to the R^2 of the model not including screen time, this model seems to perform worse.

```{r}
STComp_model_IM <- lm(Competence_Score ~ Interaction_Score + Speaking_Score + Speak_Hours_Score + Use_French_Score + French_Hours_Score + Host_French + Host_English + Phone_Language + Phone_Keyboard + Class_Level + Year_2024 + Year_2025 + Screen_Time, data = WorkingSTData)
summary(STComp_model_IM)
```
After breaking apart the Immersion Score variables, we can see that Interaction Score has significance. Additionally, Phone_language does now as well. This is also a higher R^2 than the Comp_IM model above that does not include Screen Time

# Model that includes Screen Time for y = Continue French Score
```{r}
STInt_model <- lm(Continue_French_Score ~ Immersion_Score + Host_French + Host_English + Phone_Language + Phone_Keyboard + Class_Level + Year_2024 + Year_2025 + Screen_Time, data = WorkingSTData)
summary(STInt_model)
```
Another very low R^2 value, and a global p-value that does not reject. Let's see if breaking up the significant Immersion_Score variable helps


```{r}
STInt_model_IM <- lm(Continue_French_Score ~ Interaction_Score + Speaking_Score + Speak_Hours_Score + Use_French_Score + French_Hours_Score + Host_French + Host_English + Phone_Language + Phone_Keyboard + Class_Level + Year_2024 + Year_2025 + Screen_Time, data = WorkingSTData)
summary(STInt_model_IM)
```
This did not improve the model much, still a large global p-value.

# Normality of residuals 
Chose the model with the better R^2 to check between the overall Immersion Score vs the broken up Immersion Score

# Model that does not include Screen Time for y = Competence Score
```{r}
residC <- residuals(Comp_model)
ggplot(data.frame(residC), aes(sample = residC)) + stat_qq() + stat_qq_line()
```

# Model that does not include Screen Time for y = Intention Score
```{r}
residI <- residuals(Int_model_IM)
ggplot(data.frame(residI), aes(sample = residI)) + stat_qq() + stat_qq_line()
```


# Model that includes Screen Time for y = Competence Score
```{r}
residSTC <- residuals(STComp_model_IM)
ggplot(data.frame(residSTC), aes(sample = residSTC)) + stat_qq() + stat_qq_line()
```

# Model that includes Screen Time for y = Intention Score
```{r}
residSTI <- residuals(STInt_model_IM)
ggplot(data.frame(residSTI), aes(sample = residSTI)) + stat_qq() + stat_qq_line()
```

# Constant Variance
# Model that does not include Screen Time for y = Competence Score
```{r}
ggplot(data.frame(fitted = fitted(Comp_model), resid = residC), aes(x = fitted, y = residC)) +
  geom_point()
```
Looks good, does not seem to have patterns or fan out or cluster too much

# Model that does not include Screen Time for y = Intention Score
```{r}
ggplot(data.frame(fitted = fitted(Int_model_IM), resid = residI), aes(x = fitted, y = residI)) +
  geom_point()
```
Also looks good

# Model that includes Screen Time for y = Competence Score
```{r}
ggplot(data.frame(fitted = fitted(STComp_model_IM), resid = residSTC), aes(x = fitted, y = residSTC)) +
  geom_point()
```
Also looks good

# Model that includes Screen Time for y = Intention Score
```{r}
ggplot(data.frame(fitted = fitted(STInt_model_IM), resid = residSTI), aes(x = fitted, y = residSTI)) +
  geom_point()
```
This one does show a bit of fanning out across x

## Next steps possibly looking at Interaction more as it came up as significant


