---
title: "French Study Abroad Project Analysis"
format: pdf
editor: visual
---

```{r}
library(ggplot2)
library(MASS)
library(dplyr)
```

# Load Data

```{r}
# Load the preprocessed data
french_data <- read_csv("processed_french_data.csv")

french_data
```

# EDA

## Tables

### Table of Living Arrangements and what proportion of the total n each are, as well as what proportion of the total Competence Score they are

```{r}
french_data |> group_by(Host_Family) |> summarize(n = n(), comp = sum(Competence_Score)) |> mutate(prop_count = n/sum(n), prop_comp = comp/sum(comp))
```

Top row is Student dorms, and this row seems to be the majority of student living situations.

### Spread of data for Full dataset

```{r}
french_data |> lapply(summary)
```

### Average scores grouped by our Class level covariate

```{r}
french_data |> group_by(Class_Level) |> summarise(count=n(), avgcompscore = mean(Competence_Score), avgintscore = mean(Intention_Score), .groups = "drop")
```

Seems as though by class level, with 4 being the most difficult, competence scores do increase as one would expect.

## Graphs of variables

### Boxplot of screen time

```{r}
ggplot(french_data, aes(x= Screen_Time)) + geom_boxplot(fill = "orchid")
```

Looks to be somewhat skewed, also a couple outliers to the far right

### Histogram of Competence Score, normal curve overlayed

```{r}
ggplot(french_data, aes(x= Competence_Score)) + geom_histogram(aes(y = after_stat(density)), bins = 35) + stat_function(fun = dnorm, args = list(mean = mean(french_data$Competence_Score), sd = sd(french_data$Competence_Score)),color = "red")
```

Not awful, bi-modal skewed to the left a bit

### Histogram of Intention Score, normal curve overlayed

```{r}
ggplot(french_data, aes(x= Intention_Score)) + geom_histogram(aes(y = after_stat(density)), bins= 35) + stat_function(fun = dnorm, args = list(mean = mean(french_data$Intention_Score), sd = sd(french_data$Intention_Score)),color = "red")
```

Multi-modal histogram, not very symmetric

### Histogram of Interaction Score, normal curve overlayed

```{r}
ggplot(french_data, aes(x= Interaction_Score)) + geom_histogram(aes(y = after_stat(density)), bins= 35) + stat_function(fun = dnorm, args = list(mean = mean(french_data$Interaction_Score), sd = sd(french_data$Interaction_Score)),color = "red")
```

### Visualizing Sum of total competence score by Living Arrangement -\> French Speaking Home

```{r}
ggplot(french_data, aes(x = factor(Host_Family), y = Competence_Score, fill = factor(Class_Level))) +
  geom_bar(stat = "summary", fun = "mean") + scale_fill_brewer(palette = "Greens") +
  labs(x = "Lived with a Host Family (0 = No, 1 = Yes)", y = "Total Competence Score") + theme_classic()
```

This bar plot shows that those that lived with a Host Family had a higher overall Competence score on average

### Plot of Interaction Score by Competence, colors are Class Level covariate

```{r}
ggplot(french_data, aes(x = Speaking_Score, y = Competence_Score, color = factor(Class_Level), shape = factor(Class_Level))) + geom_point() + labs(x = "Speaking Score", y = "Competence Score", color = "Class Level", shape = "Class Level")
```

This graph does not show a clear linear trend here

### Plot of Interaction Score by Intention, colors are Class Level covariate

```{r}
ggplot(french_data, aes(x = (Interaction_Score), y = Intention_Score, color = factor(Class_Level), shape = factor(Class_Level))) + geom_point()
```

Similar here, no obvious linear trend

### Screen Time Plotted by Competence Score

```{r}
ggplot(french_data, aes(x = (Screen_Time), y = Competence_Score, color = factor(Class_Level), shape = factor(Class_Level) )) + geom_point()
```

Similar results as those above

# Linear Models

The Screen Time variable had several issues with data quality and was not significant in either model, therefore we decided to remove it.

-   Screen Time does not have statistical significance in this model. However once again Immersion Score does. We also see that the global F test does not reject either, and with an adjusted R\^2 of .27, we cannot conclude much based on this model. Especially comparing this R\^2 to the R\^2 of the model not including screen time, this model seems to perform worse.

-   Also our client agreed that the variable didn't make sense as is and that we should remove it.

```{r}
competence_model2 <- lm(Competence_Score ~ Year_2024 + Year_2025 + Class_Level +
                         Host_Family + Phone_Language +
                         Phone_Keyboard + Interaction_Score + 
                         Speaking_Score + Speak_Hours_Score +
                         Use_French_Score + French_Hours_Score,
                       data = french_data)

summary(competence_model2)
```

```{r}
intention_model2 <- lm(Intention_Score ~ Year_2024 + Year_2025 + 
                         Class_Level + Host_Family + 
                         Phone_Language + Phone_Keyboard +
                         Interaction_Score + Speaking_Score + Speak_Hours_Score +
                         Use_French_Score + French_Hours_Score,
                       data = french_data)

summary(intention_model2)
```

# Model Checking using AIC

```{r}
stepAIC(competence_model2, method = "backward")
```

```{r}
# After model checking
competence_model4 <- lm(Competence_Score ~ 
                         Class_Level +
                         Speaking_Score,
                       data = french_data)

summary(competence_model4)
```

```{r}
# Competence Might Be slightly left skewed
plot(density(french_data$Competence_Score))
```

```{r}
# Residuals vs Fitted
plot(competence_model4, which = 1)
```

Looks good, does not seem to have patterns or fan out or cluster too much. Some outliers.

Checking the outliers revealed that 
- for entry 45, the student had a higher perceived Competence score (82) than usual based on their fairly low Speaking Score (10) and Class Level (1)
- For entry 16, the student had a lower perceived Competence score (29) than usual based on their fairly low Speaking Score (22) and Class Level (1)
- For entry 19, the student had a lower perceived Competence score (37) than usual based on their fairly low Speaking Score (18) and Class Level (2)


# testing Competence Score for normality
```{r}
# QQ Plot
plot(competence_model4, which = 2)
```


```{r}
shapiro.test(resid(competence_model4))
```


```{r}
stepAIC(intention_model2, method = "backward")
```

```{r}
# After model checking
intention_model4 <- lm(Intention_Score ~
                            Class_Level + 
                            Host_Family + 
                            Phone_Language + 
                            Interaction_Score +   
                            Speak_Hours_Score + 
                            Use_French_Score,
                       data = french_data)

summary(intention_model4)
```

```{r}
# Intention Score is closer to a normal or at least symmetric distribution
plot(density(french_data$Intention_Score))
```

```{r}
# Residuals vs Fitted 
plot(intention_model4, which = 1)
```

once again, we should check the noted outliers

- for entry 24, the student had a higher Intention score (18) than usual based on their other entries in the model
(class level 1, did not live with a french speaking family, did not have french as a phone language, above average interaction score etc)
- For entry 3, the student had a lower Intention score (6) than usual based on their other entries in the model
(class level 0, did not live with a french speaking family, did not french as a phone language, above average interaction score etc )
- For entry 47, the student had a lower Intention score (14) than usual based on their other entries in the model
(class level 4, lived with a french speaking family, had french as a phone language, above average interaction score etc)

# checking Intention Score for normality
```{r}
# QQ Plot
plot(intention_model4, which = 2)
```
Looks fairly normally distributed

```{r}
shapiro.test(resid(intention_model4))
```
pval > 0.05, we can assume normality


Let's look at plots of Interaction_Score, Speaking_Score, Speak_Hours_Score, Use_French_Score, and French_Hours_Score vs Competence_Score and Intention_Score

```{r}
# Something interesting I found is that Competence and Intention have 
# a moderately positive correlation 
ggplot(french_data, aes(Competence_Score, Intention_Score)) +
  geom_point() + 
  geom_smooth(method="lm", formula = y ~ x)
```

```{r}
french_data |>
  dplyr::select(everything(), -Screen_Time) |>
  cor(use = "complete.obs")
```

Ignoring the variables that are a composite of other variables, none of the variables are too highly correlated to the point that multicollinearity would be an issue. We can see that Read_French_Score, Write_French_Score, Listen_French_Score, and Speak_French_Score all have moderate to strong positive correlations with one another. Interestingly enough Competence_Score and Intention_Score almost had only positive correlations with all of the other variables, albeit many of those correlations are weak.

Might need to standardize the variables first, but for now it's probably fine
