---
title: "Cook's Distance"
output: html_notebook
---

# Competence Score Model

```{r}
cooksdc <- cooks.distance(competence_model4)
plot(cooksdc, xlab = "Observation Number", ylab = "Cook's Distance")
abline(h = 4/nrow(french_data), col = "red", lty = 2)
inflabc <- which(cooksdc > 4/nrow(french_data))
text(inflabc, cooksdc[inflabc], labels = inflabc, pos=2)
```
Looks like observation 45 is not an influential point

remove and rerun model to check
```{r}
no_outlier_comp <- french_data %>% filter(row_number() != 45)

competence_model_cd <- lm(Competence_Score ~ Year_2024 + Year_2025 + Class_Level +
                         Host_Family + Phone_Language +
                         Phone_Keyboard + Interaction_Score + 
                         Speaking_Score + Speak_Hours_Score +
                         Use_French_Score + French_Hours_Score,
                       data = no_outlier_comp)

summary(competence_model_cd)
```
Slightly changes outcome, next rerun StepAIC to see if there is a new best model

```{r}
stepAIC(competence_model_cd, method = "backward")
```

There is a new best model, run this and use R^2 to compare which model is "better"

```{r}
competence_model_cd2 <- lm(Competence_Score ~ Class_Level + Phone_Language + Interaction_Score + 
    Speaking_Score, data = no_outlier_comp)

summary(competence_model_cd2)
```
This model has a better R^2, model performs better once the influential point is removed.


## Intention Model
```{r}
cooksdi <- cooks.distance(intention_model4)
plot(cooksdi, xlab = "Observation Number", ylab = "Cook's Distance")
abline(h = 4/nrow(french_data), col = "red", lty = 2)
inflabi <- which(cooksdi > 4/nrow(french_data))
text(inflabi, cooksdi[inflabi], labels = inflabi, pos=2)
```
There are three influential points. Rerun the model and see if they improve the model once removed


```{r}
no_outlier_int <- french_data %>% filter(!row_number() %in% c(3,7,47))

intention_model_cd <- lm(Intention_Score ~ Year_2024 + Year_2025 + 
                         Class_Level + Host_Family + 
                         Phone_Language + Phone_Keyboard +
                         Interaction_Score + Speaking_Score + Speak_Hours_Score +
                         Use_French_Score + French_Hours_Score,
                       data = no_outlier_int)

summary(intention_model_cd)
```
Once again we see some changes in results after removing the influential points

```{r}
stepAIC(intention_model_cd)
```
The overall models are once again slightly different

```{r}
intention_model_cd2 <- lm(Intention_Score ~
                            Class_Level + 
                            Host_Family + 
                            Phone_Language + 
                            Phone_Keyboard +
                            Interaction_Score +   
                            Speak_Hours_Score + 
                            Use_French_Score,
                       data = no_outlier_int)

summary(intention_model_cd2)
```
The R^2 improved over 10% after removing the influential points. 

```{r}
plot(intention_model_cd2, which = 2)
```


## Normality was violated for Competence Score
```{r}
Trans_comp <- (french_data$Competence_Score)^2

competence_model5 <- lm(Trans_comp ~ 
                         Class_Level +
                         Speaking_Score,
                       data = french_data)

summary(competence_model5)
```

```{r}
plot(density(Trans_comp))
```

```{r}
plot(competence_model5, which = 2)
```


```{r}
shapiro.test(resid(competence_model5))
```
pvalue > 0.05

residuals are now normal, just heavy-tailed.

```{r}
plot(competence_model5, which = 1)
```
This plot looks about the same as the not-transformed one. 
