---
title: "EDA"
format: pdf
editor: visual
---

# french_data is the dataset that includes screen time
```{r}
french_data <- french_data |> mutate(Immersion_Score = rowSums(across(c(Interaction_Score, Speaking_Score, Speak_Hours_Score, Use_French_Score, French_Hours_Score))))
WorkingSTData <- french_data |> drop_na(8) 
```

# EDA 
## Tables
### Table of Living Arrangements and what proportion of the total n each are, as well as what proportion of the total Competence Score they are
```{r}
french_data |> group_by(Host_French, Host_English) |> summarize(n = n(), comp = sum(Competence_Score)) |> mutate(prop_count = n/sum(n), prop_comp = comp/sum(comp))
```
Top row is Student dorms, and this row seems to be the majority of student living situations.

# Spread of data for Full dataset that does not include Screen Time
```{r}
french_data[,c(18:20)] |> lapply(summary)
```

# Spread of data for Full dataset that does not include Screen Time to compare
```{r}
WorkingSTData[,c(8,18:20)] |> lapply(summary)
```
At a glance, the means do not look to be incredibly different after adding in screen time which removes 33 observations

# Average scores grouped by our Class level covariate
```{r}
french_data |> group_by(Class_Level) |> summarise(count=n(), avgcompscore = mean(Competence_Score), avgintscore = mean(Continue_French_Score), avgImmscore = mean(Immersion_Score), .groups = "drop")
```
Seems as though by class level, with 4 being the most difficult, competence scores do increase as one would expect.


## Graphs of variables

### Boxplot of screen time
```{r}
ggplot(WorkingSTData, aes(x= Screen_Time)) + geom_boxplot(fill = "orchid")
```

# Histogram of Competence Score, normal curve overlayed
```{r}
ggplot(french_data, aes(x= Competence_Score)) + geom_histogram(aes(y = after_stat(density)), bins = 35) + stat_function(fun = dnorm, args = list(mean = mean(french_data$Competence_Score), sd = sd(french_data$Competence_Score)),color = "red")
```


# Histogram of Intention Score, normal curve overlayed
```{r}
ggplot(french_data, aes(x= Continue_French_Score)) + geom_histogram(aes(y = after_stat(density)), bins= 35) + stat_function(fun = dnorm, args = list(mean = mean(french_data$Continue_French_Score), sd = sd(french_data$Continue_French_Score)),color = "red")
```

# Histogram of Immersion Score, normal curve overlayed
```{r}
ggplot(french_data, aes(x= Immersion_Score)) + geom_histogram(aes(y = after_stat(density)), bins= 35) + stat_function(fun = dnorm, args = list(mean = mean(french_data$Immersion_Score), sd = sd(french_data$Immersion_Score)),color = "red")
```


# Visualizing Sum of total competence score by Living Arrangement -> French Speaking Home
```{r}
ggplot(french_data, aes(x = factor(Host_French), y = Competence_Score, fill = factor(Class_Level))) +
  geom_bar(stat = "summary", fun = "mean") + scale_fill_brewer(palette = "Reds") +
  labs(x = "Lived in a French Speaking Home (0 = No, 1 = Yes)", y = "Total Competence Score") + theme_classic()
```
This bar plot shows that those that lived in a French Speaking home had a higher overall Competence score on average


#Plot of Immersion Score by Competence, colors are Class Level covariate
```{r}
ggplot(french_data, aes(x = (Immersion_Score), y = Competence_Score, color = factor(Class_Level))) + geom_point() + geom_abline()
```
This graph does not show a clear linear trend here

#Plot of Immersion Score by Intention, colors are Class Level covariate
```{r}
ggplot(french_data, aes(x = (Immersion_Score), y = Continue_French_Score, color = factor(Class_Level))) + geom_point()
```
Similar here, no obvious linear trend

# Screen Time Plotted by Competence Score
```{r}
ggplot(WorkingSTData, aes(x = (Screen_Time), y = Competence_Score, color = factor(Class_Level))) + geom_point()
```
Similar results as those above
