---
title: "ST542 French Immersion Project"
format: pdf
editor: visual
---

## Data Preprocessing

Convert ordinal categorical data to numbers.

Add up subgroups to create one column per variable.

```{r}
# Load libraries
library(tidyverse)
library(readxl)
```

```{r}
french_post_survey <- read_excel("French_Post_Survey.xlsx",sheet=1,
                      col_names = c("Timestamp", "Email", "Changed_Major",
                                   "Changed_Minor", "Masters", "Feelings",
                                   "Feelings_Factors", "Class_Level", 
                                   "Living_Arrangement", "LA_a1",
                                   "LA_a2", "LA_a3", "LA_a4", "LA_b1", "LA_b2",
                                   "LA_b3", "Groupness1", "Groupness2", 
                                   "Groupness3", "Groupness4", "Groupness5",
                                   "Phone_Language", "Phone_Keyboard",
                                   "Screen_Time", "Interact1", "Interact2",
                                   "Interact3", "Interact4", "Interact5",
                                   "Interact6", "Interact7", "Interact8",
                                   "Interact9", "Interact10", "Speaking1",
                                   "Speaking2", "Speaking3", "Speaking4", "Speaking5",
                                   "Speak_Hours1", "Speak_Hours2",
                                   "Speak_Hours3", "Speak_Hours4", 
                                   "Speak_Hours5", "Use_French1", "Use_French2",
                                   "Use_French3", "Use_French4", 
                                   "French_Hours1", "French_Hours2", 
                                   "French_Hours3", "French_Hours4",
                                   "Read_French1", "Read_French2", 
                                   "Read_French3", "Read_French4", 
                                   "Read_French5", "Write_French1",
                                   "Write_French2", "Write_French3",
                                   "Write_French4", "Write_French5",
                                   "Listen_French1", "Listen_French2",
                                   "Listen_French3", "Listen_French4",
                                   "Listen_French5", "Speak_French1",
                                   "Speak_French2", "Speak_French3", 
                                   "Speak_French4", "Speak_French5",
                                   "Continue_French1", "Continue_French2",
                                   "Continue_French3", "Continue_French4",
                                   "Continue_French5", "Code", "Extra1", 
                                   "Extra2", "Extra3"),
                      skip = 1)

# The last 9 values were from extra work, not survey responses
french_post_survey <- french_post_survey[1:61,]
french_post_survey
```

```{r}
# Subsetting Columns We Will Use In Our Analysis
french_data <- french_post_survey |>
  select("Timestamp", "Class_Level", "Living_Arrangement", "LA_a2", "LA_b1",
         "Phone_Language", "Phone_Keyboard", "Screen_Time", "Interact1", "Interact2",
         "Interact3", "Interact4", "Interact5", "Interact6", "Interact7", "Interact8",
         "Interact9", "Interact10", "Speaking1", "Speaking2", "Speaking3", 
         "Speaking4", "Speaking5", "Speak_Hours1", "Speak_Hours2", 
         "Speak_Hours3", "Speak_Hours4", "Speak_Hours5", 
         "Use_French1", "Use_French2", "Use_French3", "Use_French4", "French_Hours1", 
         "French_Hours2", "French_Hours3", "French_Hours4", "Read_French1", 
         "Read_French2", "Read_French3", "Read_French4", "Read_French5", 
         "Write_French1", "Write_French2", "Write_French3", "Write_French4", 
         "Write_French5", "Listen_French1", "Listen_French2", "Listen_French3", 
         "Listen_French4", "Listen_French5", "Speak_French1", "Speak_French2", 
         "Speak_French3", "Speak_French4", "Speak_French5", "Continue_French1", 
         "Continue_French2", "Continue_French3", "Continue_French4", "Continue_French5")

french_data
```

### Preliminary Data Processing

The Screen Time variable has been calculated in another excel file and needs to be added to this dataset.

```{r}
screen_time_data <- read_excel("ScreenTimeDat.xlsx")

french_data$Screen_Time <- screen_time_data$AvgScreenTime

french_data
```

In the class level variable there are FLF 201 - Intermediate French 1 and FLF 202 - Intermediate French 2. These two however are incorrectly entered data, so these values will have to be removed before we begin processing the rest of the data.

```{r}
french_data <- french_data |>
  filter(Class_Level != "FLF 201",
         Class_Level != "FLF 202")

french_data
```

```{r}
french_data |>
  group_by(Class_Level) |>
  summarize(count = n())
```

### Response Variables

Reading, Writing, Listening, and Speaking French Perceived Skill. Intention to continue learning French in some capacity.

```{r}
french_data |>
  group_by(Read_French2) |>
  summarize(count = n())
```

This is typical ordinal categorical data. Need to convert it to a 1 to 5 scale.

```{r}
# To convert ordinal categorical values to numbers
ordinal_dict <- c("Not sure" = 1,
                  "Somewhat sure" = 2,
                  "Moderately sure" = 3,
                  "Very sure" = 4, 
                  "Completely sure" = 5)

ordinal_columns <- c("Read_French1", "Read_French2", "Read_French3",
                     "Read_French4", "Read_French5", "Write_French1",
                     "Write_French2", "Write_French3", "Write_French4",
                     "Write_French5", "Listen_French1", "Listen_French2",
                     "Listen_French3", "Listen_French4", "Listen_French5",
                     "Speak_French1", "Speak_French2", "Speak_French3",
                     "Speak_French4", "Speak_French5", "Continue_French1",
                     "Continue_French2", "Continue_French3", "Continue_French4",
                     "Continue_French5")

convert_ordinal <- function(col) {
  return(ordinal_dict[col])
}
```

```{r}
# How to convert ordinal categorical data to numerical
french_data <- french_data |>
  mutate(across(all_of(ordinal_columns), convert_ordinal))

french_data
```

### Hours Speaking or Using French

```{r}
# Lets look at another type of ordinal column
french_data |>
  group_by(Speak_Hours1) |>
  summarize(count = n())
```

```{r}
# To convert ordinal categorical hour values to numbers
ordinal_hours_dict <- c("0-1" = 0,
                        "1-2" = 1,
                        "2-3" = 2,
                        "3-4" = 3, 
                        "4-5" = 4,
                        "5+" = 5)

ordinal_hours_columns <- c("Speak_Hours1", "Speak_Hours2","Speak_Hours3",
                           "Speak_Hours4", "Speak_Hours5", "French_Hours1",
                           "French_Hours2", "French_Hours3", "French_Hours4")

convert_ordinal_hours <- function(col) {
  return(ordinal_hours_dict[col])
}
```

```{r}
# How to convert ordinal categorical data to numerical
french_data <- french_data |>
  mutate(across(all_of(ordinal_hours_columns), convert_ordinal_hours))

french_data
```

### Interaction With French People Frequency

```{r}
# Lets look at another type of ordinal column
interact<- french_data |>
  group_by(Interact10) |>
  summarize(count = n())

interact$Interact10
```

```{r}
# To convert ordinal categorical interaction values to numbers
ordinal_interact_dict <- c("not available" = 1,
                           "never" = 2,
                           "rarely (<1 times/week)" = 3,
                           "occasionally (1-2 times/week)" = 4, 
                           "frequently (3-5 times/week)" = 5)

ordinal_interact_columns <- c("Interact1", "Interact2", "Interact3", 
                              "Interact4", "Interact5", "Interact6",
                              "Interact7", "Interact8", "Interact9",
                              "Interact10")

convert_ordinal_interact <- function(col) {
  return(ordinal_interact_dict[col])
}
```

```{r}
# How to convert ordinal categorical data to numerical
french_data <- french_data |>
  mutate(across(all_of(ordinal_interact_columns), convert_ordinal_interact))

french_data
```

### Language on Phone and Keyboard Language on Phone

"Phone_Language", "Phone_Keyboard" We may or may not use these.

```{r}
# Lets look at the values in Language on Phone and Keyboard Language on Phone
french_data |>
  group_by(Phone_Language) |>
  summarize(count = n())
```

```{r}
french_data |>
  group_by(Phone_Keyboard) |>
  summarize(count = n())
```

Looks like the general trend for both of these columns is that students either did or did not use French on their phone. We can convert these into binary variables based on this.

```{r}
# Convert to binary where 0 represents no French on phone, and 1 represents French on phone
binary_phone_dict <- c("English" = 0,
                       "English, French" = 1,
                       "English, French, German" = 1,
                       "English, French, Korean" = 1,
                       "English, French, Spanish" = 1,
                       "English, French, spanish" = 1,
                       "English, French, Spanish, Greek" = 1,
                       "English, French, Spanish, Korean" = 1,
                       "English, French, Spanish, Portuguese" = 1,
                       "English, French, spanish" = 1,
                       "French" = 1,
                       "German" = 0)


binary_phone_columns <- c("Phone_Language", "Phone_Keyboard")

convert_binary_phone <- function(col) {
  return(binary_phone_dict[col])
}
```

```{r}
# Convert the phone variables to binary
french_data <- french_data |>
  mutate(across(all_of(binary_phone_columns), convert_binary_phone))

french_data
```

```{r}
french_data |>
  group_by(Phone_Language) |>
  summarize(count = n())
```

```{r}
french_data |>
  group_by(Phone_Keyboard) |>
  summarize(count = n())
```

### Living Arrangements

"Living_Arrangement" Has it's own encoding.

"LA_a1" may not use - Doesn't have a consistent entry.

"LA_a3", "LA_b2", "LA_b3" all do not have enough yes answers. "LA_a4" divides the data into too small of groups and doesn't really seem all that important.

"LA_a2", "LA_b1" All yes or no.

```{r}
french_data |>
  group_by(Living_Arrangement) |>
  summarize(count = n())
```

There are only two people who lived alone in a room, apartment or house. There's only one person who lived in those environments with others who were not French speakers. I think it's reasonable to say these are equivalent to students who lived in a dorm alone or lived in a dorm with roommates.

Converting this to just Host_French and Host_English, with the base case where both are 0 being Student_Housing.

```{r}
french_data <- french_data |>
  mutate(Host_French = if_else(startsWith(Living_Arrangement, "a"), 
                               if_else(LA_a2 == "Yes", 0, 1),
                               0),
         Host_English = if_else(startsWith(Living_Arrangement, "a"), 
                                if_else(LA_a2 == "Yes", 1, 0), 
                                0),
         Host = if_else(startsWith(Living_Arrangement, "a"),
                        1, 0)) |>
  select(Timestamp, Class_Level, Host, Host_French, Host_English,
         everything(), -Living_Arrangement, -LA_a2, -LA_b1)

french_data
```

```{r}
french_data |>
  group_by(Host_French) |>
  summarize(count = n())
```

```{r}
french_data |>
  group_by(Host_English) |>
  summarize(count = n())
```

```{r}
french_data |>
  group_by(Host) |>
  summarize(count = n())
```

### Class Level

"Class_Level" Need to make sure we know which is the highest.

Language levels go from A1 - beginner, A2 - Elementary, B1 - Lower Intermediate, B2 - Upper Intermediate, C1 - Advanced, C2 - Mastery.

```{r}
french_data |>
  group_by(Class_Level) |>
  summarize(count = n())
```

```{r}
# Convert ordinal Class Level
class_level_dict <- c("Summer Study Abroad A1-level language class" = 0,
                      "Summer Study Abroad A2-level language class" = 1,
                      "Summer Study Abroad B1-level language class" = 2,
                      "Summer Study Abroad B2-level language class" = 3,
                      "Summer Study Abroad C1-level language class" = 4)


class_level_column <- c("Class_Level")

convert_class_level <- function(col) {
  return(class_level_dict[col])
}
```

```{r}
# Convert the class level variable
french_data <- french_data |>
  mutate(across(all_of(class_level_column), convert_class_level))

french_data
```

### Year

"Timestamp" Need to convert to just the year.

Year is being treated as categorical with two binary variables one for 2024 and the other 2025, in which case if both are 0, then we have 2023.

```{r}
# Convert the class level variable
french_data <- french_data |>
  mutate(Year = year(Timestamp),
         Year_2024 = if_else(Year == 2024, 1, 0),
         Year_2025 = if_else(Year == 2025, 1, 0)) |>
  select(Year_2024, Year_2025, everything(), -Timestamp, -Year)

french_data
```

```{r}
french_data |>
  group_by(Year_2024) |>
  summarize(count = n())
```

```{r}
french_data |>
  group_by(Year_2025) |>
  summarize(count = n())
```

### Scores

Might make sense to still look at the Read, Write, Listen, and Speak scores individually so I included them.

```{r}
french_data <- french_data |>
  mutate(Interaction_Score = rowSums(across(starts_with("Interact"))),
         Speaking_Score = rowSums(across(starts_with("Speaking"))),
         Speak_Hours_Score = rowSums(across(starts_with("Speak_Hours"))),
         Use_French_Score = rowSums(across(starts_with("Use_French"))),
         French_Hours_Score = rowSums(across(starts_with("French_Hours"))),
         Read_French_Score = rowSums(across(starts_with("Read_French"))),
         Write_French_Score = rowSums(across(starts_with("Write_French"))),
         Listen_French_Score = rowSums(across(starts_with("Listen_French"))),
         Speak_French_Score = rowSums(across(starts_with("Speak_French"))),
         Competence_Score = rowSums(across(c(Read_French_Score,
                                             Write_French_Score,
                                             Listen_French_Score,
                                             Speak_French_Score))),
         Continue_French_Score = rowSums(across(starts_with("Continue_French")))) |>
  select(Year_2024, Year_2025, Class_Level, Host, Host_French, Host_English,
         Phone_Language, Phone_Keyboard, Screen_Time, Interaction_Score,
         Speaking_Score, Speak_Hours_Score, Use_French_Score, French_Hours_Score,
         Read_French_Score, Write_French_Score, Listen_French_Score,
         Speak_French_Score, Competence_Score, Continue_French_Score) 

french_data
```

### Final Export

```{r}
write_csv(french_data, "processed_french_data.csv")
```
